---
title: "UAS Body Condition: Comparison of Altitude/Range, Pitch and Roll from Drone and LRF Values"
# author: "Stacie Hardy"
date: "Date Created: `r format(Sys.time(), '%m/%d/%Y')`"
output: pdf_document

knit: (function(inputFile, encoding) { 
  rmarkdown::render(
    inputFile, 
    encoding = encoding, 
    output_file = file.path(dirname(inputFile), 'DataSummary_Dyson_20220422_fl01.pdf')) 
  })

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) 

library(tidyverse)
library(stringr)
library(lubridate)
library(RPostgreSQL)
library(sf)

## IMPORTANT: SET VARIABLES!!!!
flight_date <- '2022-04-22'
flight_number <- 'fl01'

# Connect to DB and get starting data
con <- RPostgreSQL::dbConnect(PostgreSQL(), 
                              dbname = Sys.getenv("pep_db"), 
                              host = Sys.getenv("pep_ip"), 
                              port = Sys.getenv("pep_port"),
                              user = Sys.getenv("pep_admin"), 
                              password = Sys.getenv("admin_pw"))

# Get data from DB
images <- RPostgreSQL::dbGetQuery(con, paste("SELECT * FROM body_condition.summ4report_images WHERE flight_date_local = \'", flight_date, "\' AND flight_number = \'", flight_number, "\'", sep = ""))

lrf_name <- unique(images$pep_id)

images.sf <- sf::st_as_sf(images, coords = c("exif_longitude", "exif_latitude"), crs = st_crs(4326))

# Import LRF data
lrf <- RPostgreSQL::dbGetQuery(con, paste("SELECT * FROM body_condition.summ4report_lrf WHERE flight_date_local = \'", flight_date, "\' AND flight_number = \'", flight_number, "\'", sep = "")) %>%
  filter(gps_dt > as.Date(flight_date_local) - 1)

lrf.sf <- lrf %>% filter(gps_latitude != -99)
lrf.sf <- sf::st_as_sf(lrf.sf, coords = c("gps_longitude", "gps_latitude"), crs = st_crs(4326))
```
## Introduction
The purpose of this document is to present a comparison of data streams recorded by the drone and by the associated laser rangefinder during a UAS flight. The data in this report were collected on **`r flight_date`** on **flight `r flight_number`**. Images were collected using the **`r unique(images$lrf)` drone**, and laser rangefinder (LRF) data were collected using **LRF `r unique(images$pep_id)` (`r unique(images$aircraft_make)` `r unique(images$aircraft_model)`)**. Exif data are only available when images were taken, and up to 5 values were recorded per second by the LRF. The data presented here are reflective of the timestamp recorded to the image's exif data, and an offset of **`r round(unique(images$exif2gps_offset_sec) / 60, 2)` minutes** applied to the LRF data.

## Comparison of Location Data
```{r location, echo = FALSE, fig.height=4}
ggplot() +
  geom_sf(data = lrf.sf, color = 'sienna3', size = 0.5) +
  geom_sf(data = images.sf, fill = 'chartreuse3', color = 'chartreuse3', size = 1, shape = 24) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```

**Figure 1. LRF GPS location compared to GPS location from image exif.** The locations of the LRF GPS are depicted as orange dot, and the locations where images were taken by the drone are depicted as green triangles.  

\newpage

## Comparison of Altitude Values

```{r altitude, echo=FALSE}
ggplot() + 
  geom_point(data = lrf %>% mutate(gps_dt = gps_dt), aes(gps_dt, laser_range_raw_m), colour = "sienna3", size = 0.2) + 
  geom_point(data = images, aes(adjusted_image_dt, exif_relative_altitude_m), colour = "chartreuse3", size = 0.2) + 
  geom_point(data = images, aes(adjusted_image_dt, exif_altitude_m), colour = "navy", size = 0.2) + 
  xlim(min(lrf$gps_dt), max(lrf$gps_dt)) +
  labs(x = "Date/time", y = "Relative altitude / LRF range (m)")
```

**Figure 2. LRF range (m) compared to relative altitude (m) from image exif.** The range recorded by the laser rangefinder is displayed as orange dots, the relative altitude recorded to each image is displayed as a green dot, and the GPS altitude recorded to each image is displayed as a blue dot.      

\newpage

## Comparison of Pitch Values

```{r pitch, echo=FALSE}
ggplot() + 
  geom_line(data = lrf %>% mutate(gps_dt = gps_dt), aes(gps_dt, imu_pitch), colour = "sienna3") + 
  geom_point(data = images, aes(adjusted_image_dt, exif_gimbal_pitch), colour = "chartreuse3", size = 0.2) +
  geom_point(data = images, aes(adjusted_image_dt, exif_pitch), colour = "navy", size = 0.2) + 
  xlim(min(lrf$gps_dt), max(lrf$gps_dt)) +
  labs(x = "Date/time", y = "Pitch (degrees)")

```

**Figure 3. LRF pitch (degrees) compared to pitch (degrees) from image exif.** The pitch recorded by the laser rangefinder is displayed as orange dots, the gimbal pitch recorded to each image is displayed as a green dot, and the drone pitch recorded to each image is displayed as a blue dot.   

\newpage 

## Comparison of Roll Value

```{r roll, echo=FALSE}
ggplot() + 
  geom_line(data = lrf %>% mutate(gps_dt = gps_dt), aes(gps_dt, imu_roll), colour = "sienna3") + 
  geom_point(data = images, aes(adjusted_image_dt, exif_gimbal_roll), colour = "chartreuse3", size = 0.2) +
  geom_point(data = images, aes(adjusted_image_dt, exif_roll), colour = "navy", size = 0.2) + 
  xlim(min(lrf$gps_dt), max(lrf$gps_dt)) +
  labs(x = "Date/time", y = "Roll (degrees)")
```

**Figure 4. LRF roll (degrees) compared to roll (degrees) from image exif.** The roll recorded by the laser rangefinder is displayed as orange dots, the gimbal roll recorded to each image is displayed as a green dot, and the drone roll recorded to each image is displayed as a blue dot.